Enhanced Material Point Method to face dynamic problems:\thanks{Funding: This study was funded by Agustín de Betancourt Foundation (grant number 262390106114).}

### 2020-03-25 main.tex
Versión día 25-3-2020

\usepackage[lofdepth,lotdepth]{subfig}
% \usepackage{graphicx}
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

\maketitle
\begin{abstract}
  This document is devoted to describe the suitability of the Local
  \textit{maximum-entropy} (LME) meshfree approximation technique under the framework of the
  Material Point Method for dynamic problems. 
  \keywords{LME \and MPM \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Since the proposal of the Material Point Method (\textbf{MPM}) by
Sulsky {\it  et al.} (1994)\cite{Sulsky1994} as a generalization to
solids of the fluid implicit particle (FLIP) method
\cite{Brackbill1986}. It popularity has increased due to its ability
to deal with large strain regime without suffer mesh distortion
inaccuracies. However, this method suffers other kind of
instabilities, such those when material points crossing cell
boundaries. This give rise to the development of other interpolation
techniques to overcome this limitation such as the generalized
interpolation material point method (GIMP) Bardenhagen \& Kober
(2004)\cite{Bardenhagen2004}, the dual
domain material point method (DDMP) Zhang {\it et al.}
(2011)\cite{Zhang2011a}, the B-Spline MPM Tran {\it et al.}
(2019)\cite{Tran2019a} or the Conservative Taylor
Least Squares reconstruction Wobbes {\it et al.}
(2018)\cite{E_Wobbes_2018}. This techniques are devoted to mitigate the
``grid crossing'' error.
Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using techniques \cite{Tran2019e}. These numerical
inaccuracies also known as wiggles are due to the time discretization
technique. A simple approach to face those spurious noises is to add a
damping source to the equilibrium equations. To overcome this
limitation, many researchers has proposed alternative time integration
schemes which reduce them by filtering high-frequencies. Some of these
attempts are the implicit GIMP (iGIMP) Charlton {\it et al.}
(2017)\nocite{Charlton2017} or the generalised-$\alpha$ scheme for the
MPM proposed by Tran \& Solowski (2019)\nocite{Tran2019e}.

The aim of this paper is to mitigate this spurious oscillations by the
employ of a suitable combination of the maximum-entropy (or local
\textit{max-ent}) $\gamma$ family shape functions, and the proposal of a explicit 
predictor-corrector scheme. Local \textit{max-ent} shape functions was
first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite elements (\textbf{FE}) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness of the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions \cite{Arroyo2006}. Furthermore, as
it is not based in a voxel discretization as the GIMP, it is no prone to suffer
voxel domains overlap, which can introduce severe inaccuracies. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)\cite{Navas2018a} and Li {\it et al.} (2012)\cite{Li2012} for
Optimal Transportation Meshfree (\textbf{OTM}) method. And more
recently under the MPM framework by Wobbes {\it et
  al.}(2020)\cite{Wobbes2020} but limiting the $\gamma$ parameter to
zero. On the other hand, the time integration scheme here proposed is an accommodation of
the Newmark a-form ($\gamma = 0.5$ and $\beta = 0$)\footnote{Notice
that these $\gamma$ and $\beta$ parameters does not have the same
meaning for the LME shape functions.} which is the central difference
explicit (CD). We have choose this method among other suitable
alternatives as Wilson {\it et al.} (1972)\nocite{Wilson1972} or Chung
\& Hulbert (1993)\nocite{Geranlized_alpha_1993} because it simplicity
and it good performance dealing with solids dynamics problems like those
presented in \cite{Navas2018a}.

@@

Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$, but also the
virtual field $\vect{v}^{\psi}_{p}$. So employing the definition of the
material integral, where we recover the Riemann integral definition as
an addition  of a finite set of points. Therefore individual terms in
\eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a} \cdot \vect{v}^{\psi}} =
    \dot{\vec{v}}_{p} \cdot \vect{v}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{v}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{v}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{v}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{v}^{\psi}} = \vec{t}^s_{p} \cdot \vect{v}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Here is
where the second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated. Where $N_n$ is the number of grid
nodes. This mesh is employed as a support to compute gradients and divergences.

Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces}, 
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
and approximating the real and virtual velocity  field of the particle
$p$ as
\begin{equation}
  \label{eq:Approx-velocity-mesh}
  \vec{v}_{p} = N_I(\vec{x}_p) \vect{v}_I
  \quad \text{and} \quad \vec{v}^{\psi}_{p} = N_I(\vec{x}_p) \vect{v}_I^{\psi},
\end{equation}
\begin{equation}
  \label{eq:Approx-gradvelocity-mesh}
  \Grad{\vect{v}}_p = \vect{v}_I N_I(\vec{x}_p)
  \quad \text{and} \quad \vec{v}^{\psi}_{p} = \sum^{Nn}_I N_I(\vec{x}_p) \vect{v}_I^{\psi},
\end{equation}
we reach to the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} =
  \sum^{Np}_{p} N_{Ip} m_p N_{Jp}.
\end{equation}
To improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Later, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \sum^{N_p}_{p} \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = \sum^{N_p}_{p} N_{Ip} \vec{b}_{p} m_p  + \sum^{N_p}_{p}
  N_{Ip} \vec{t}^s_{p} m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \sum^{Nn}_I \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order
temporal integration scheme is required. Therefore, time is
discretized in to a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation it is solved, the values in the nodes are
interpolated back in to the particles and each particle it is advected
to the new position,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = \sum^{N_n}_{I}N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = \sum^{N_n}_{I}N_{Ip}\ \vec{v}_{I}  
\end{equation}
In the MPM literature, this equations
(\ref{eq:particle_balance_forces3}) and (\ref{eq:Updated_Lagrangian}),
are solved with an explicit forward Euler algorithm. The time integration scheme of the MPM has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in Figure \ref{fig:MPM_algorithm}.

\begin{figure}\sidecaption
  \centering
  \resizebox{0.55\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{MPM standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}

Other authors have proposed many others time integration alternatives
like \cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the
first publication on the MPM \cite{Sulsky1994}, the nodal acceleration
was employed to update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} point out, this algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \sum^{N_n}_{I} \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I}\frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for the MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\
  \sum^{N_n}_{I} N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + 
  \sum^{N_n}_{I}N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = \sum^{N_n}_{I}N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

This scheme has prof to damps out the higher frequency noises
\cite{Tran2019e}. But it can present the same numerical instabilities
as in \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v} when
nodal masses become infinitesimal.

@@

In this section, an explicit predictor-corrector time integration
scheme is presented. It is based in the Newmark a-form 
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}

As the MPM has a nodal stage, it is possible to apply this methods
successfully in the MPM framework as was proved by
\cite{Tran2019e}. Taking the predictor definition from the classic
literature \cite{Hughes2000} and calculating nodal velocity, and
updating particles position employing nodal values of velocity and
acceleration. Therefore we get the following expression for the
\textbf{predictor} stage for both velocity and displacement.
\begin{equation}
  \label{eq:5}
  \vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
  \underbrace{\sum^{N_n}_{I}N_{Ip}^k\ \vec{v}_{I}^{k}}_{\vec{v}_p^{k}} + \frac{1}{2}\Delta
  t^2\ \underbrace{\sum^{N_n}_{I}\frac{N_{Ip}^k\ \vec{f}_{I}^{k}}{\tens{m}_I^k}}_{\vec{a}_p^{k}}
\end{equation}
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{v}_I^{k+1} = \vec{v}_I^n + (1 - \gamma)\ \Delta t\ \vec{a}_I^k
  \quad , \quad \vec{a}_I^k = \frac{\vec{f}_{I}^{k}}{\tens{m}_I^k}
\end{equation}

>>>>

\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{subfigure} % subfiguras
%
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{epstopdf}
\usepackage{mathrsfs} % para formato de letra
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{array}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{color}
\usepackage{makeidx}
\usepackage{xstring}
\usepackage{setspace}
\usepackage{epsfig}

@@

\maketitle
\begin{abstract}
  Material Point Method (MPM) has arise in the recent years as an
  alternative to Finite Element Method (FEM) under the large
  deformation regime. However, the simulation of shock waves
  propagation and other high frequency problems is still challenging
  under this approach due the incapability of the standard MPM time
  integration scheme to filter spurious noises. To overcome this
  limitation in this paper a explicit predictor-corrector integration
  method has been proposed. Its superior performance mitigates the
  presence of spurious oscillations with minimal dissipation in high
  frequency problems. Other source of numerical noise in the MPM is
  due to the lack of smoothness of the interpolation functions, and
  occurs due to material points crossing computational grid
  boundaries. This noise results in spurious local variations at the material
  points, where strain-stress fields are computed. This could invalidate
  the solution in one case, or damage it in another. To overcome it,
  this document adopts the \textit{local maximum-entropy approximation
    schemes} (LME) a robust substitute for the wide range of shape
  function in the MPM. Local \textit{max-ent} approximation may be regarded as a
  \textit{thermalization} of Delaunay triangulation which resolves the
  degenerate cases resulting from the lack or uniqueness of the
  triangulation. Furthermore, by modifying one user-defined parameter they
  are able to behave like finite elements or as a mesh-free
  method. This capability allows to face a wide range of physics
  with a single the shape function. Finally this paper demonstrates
  the performance of both improvements thorough numerical examples.    
  \keywords{LME \and MPM \and Explicit predictor-corrector \and Dynamic problems}
  % \PACS{PACS code1 \and PACS code2 \and more}
  % \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Since the proposal of the MPM by Sulsky {\it  et al.}
(1994)\cite{Sulsky1994} as a generalization to solids of the Fluid
Implicit Particle (FLIP) method \cite{Brackbill1986}. It popularity
has increased due to its ability to deal with large strain regime
without suffer mesh distortion inaccuracies.

However, this method suffers other kind of instabilities, such those
when material points crossing cell boundaries. This give rise to the
development of other interpolation techniques to overcome this
limitation such as the generalized interpolation material point method
(GIMP) Bardenhagen \& Kober (2004)\cite{Bardenhagen2004}, which has
demonstrate to have a good performance in the finite deformation
regime. However, in the absence of a regular grid, construction of the
weighting functions is only achieved at considerable effort and
computational cost. Furthermore, as it is based in a voxel
discretization, it is prone to suffer voxel domains overlap or gaps when the
material point mesh becomes irregular, which can introduce severe
inaccuracies. This is similar to the difficulty encountered by the
finite element methods due to element distortion.
A more robust alternative is the dual domain material point method
(DDMP) proposed by Zhang {\it et al.}
(2011)\cite{Zhang2011a}. Unfortunately this method shows an
unsatisfactory behaviour when particle/cell ratio
decreases~\cite{DHAKAL2016301}, therefore DDMP requires a large number
of particles needed for convergence, this
makes the method very expensive.

the B-Spline MPM
Tran {\it et al.} (2019)\cite{Tran2019a}

the Conservative Taylor Least Squares \textbf{CTLS} reconstruction Wobbes {\it et al.}
(2018)\cite{E_Wobbes_2018}.

Local \textit{max-ent} shape functions was first introduced by Arroyo \& Ortiz (2006)\cite{Arroyo2006}, it belongs to the class of convex 
approximation schemes and provides a seamless transition between
finite elements (\textbf{FE}) and mesh-free interpolations. The
approximation scheme is based on a compromise between minimizing the
width of the shape function support and maximizing the information
entropy of the approximation. The local \textit{max-ent} approximation
may be regarded as a regularization, or \textit{thermalization}, of
Delaunay triangulation which effectively resolves the degenerate cases
resulting from the lack of uniqueness or the triangulation. Local
\textit{max-ent} basis functions possess many desirable properties for
mesh-free algorithms. First of all, they are entirely defined by the
nodal set and the domain of analysis. They are also non-negative,
satisfy the partition of unity property, and provide an exact
approximation for affine functions \cite{Arroyo2006}. This
approximation scheme has been proof to have a good performance under
the dynamic regime by other researchers like Navas {\it et al.}
(2018)\cite{Navas2018a} and Li {\it et al.} (2012)\cite{Li2012} for
Optimal Transportation Meshfree (\textbf{OTM}) method. And more
recently under the MPM framework by Wobbes {\it et
  al.}(2020)\cite{Wobbes2020} but limiting the $\gamma$ parameter to
zero.

This techniques are devoted to mitigate the
``grid crossing'' error. Nevertheless, in the presence of shock waves spurious
numerical noises appears despite of this using techniques
\cite{Tran2019e}.

These numerical inaccuracies also known as wiggles
are due to the time discretization technique. A simple approach to
face those spurious noises is to add a non-physical damping source to the
equilibrium equations. To avoid this, many researchers
has proposed alternative time integration schemes which reduce the
presence of high-frequency noises by filtering them. Some of these
attempts are the implicit GIMP (iGIMP) Charlton {\it et al.}
(2017)\nocite{Charlton2017} or the generalised-$\alpha$ scheme for the
MPM proposed by Tran \& Solowski (2019)\nocite{Tran2019e}.

On the other hand, the time integration scheme here proposed is an accommodation of
the Newmark a-form ($\gamma = 0.5$ and $\beta = 0$)\footnote{Notice
that these $\gamma$ and $\beta$ parameters does not have the same
meaning for the LME shape functions.} which is the central difference
explicit (CD). We have choose this method among other suitable
alternatives as Wilson {\it et al.} (1972)\nocite{Wilson1972} or Chung
\& Hulbert (1993)\nocite{Geranlized_alpha_1993} because it simplicity
and it good performance dealing with solids dynamics problems like those
presented in \cite{Navas2018a}.

The aim of this paper is to mitigate this spurious oscillations by the
employ of a suitable combination of the maximum-entropy (or local
\textit{max-ent}) $\gamma$ family shape functions, and the proposal of a explicit 
predictor-corrector scheme.

@@

Each material point is assigned with initial values of position,
velocity, mass, volume and stress denoted by $\vec{x}_p$,
$\vec{v}_p$, $m_p$,  $V_p$ and $\tens{\sigma}_p$, but also the
virtual field $\vect{v}^{\psi}_{p}$. So employing the definition of the
material integral, where we recover the Riemann integral definition as
an addition of a finite set of points, and their volumes are
interpreted as quadrature weights. Therefore individual terms in
\eqref{eq:BalanceMomentum_wf} are solved as follows. 
\begin{itemize}
\item Acceleration forces :
\begin{equation}
    \label{eq:particle_acceleration_forces}
    \Integral{3}{\rho\ \vec{a} \cdot \vect{v}^{\psi}} =
    \dot{\vec{v}}_{p} \cdot \vect{v}^{\psi}_{p}\ m_p.
  \end{equation}\\
\item Internal forces :
  \begin{equation}
    \label{eq:particle_internal_forces}
    \Integral{3}{\tens{\sigma}\ \colon \tens{\varepsilon}^{\psi}} =
   \tens{\sigma}_{p}\ \colon \tens{\varepsilon}^{\psi}_p \ V_p.
  \end{equation}\\
\item Body forces :
\begin{equation}
  \label{eq:particle_body_forces}
  \Integral{3}{\rho\ \vec{b} \cdot \vect{v}^{\psi} } = 
  \vec{b}_{p} \cdot \vect{v}^{\psi}_p\ m_p.
\end{equation}\\
\item Loads :
\begin{equation}
  \begin{aligned}
    \label{eq:particle_load_forces}
    \Integral{2}{\vec{t}\ \vect{v}^{\psi}} = \Integral{2}{\rho\
      \vec{t}^s \cdot \vect{v}^{\psi}} = \vec{t}^s_{p} \cdot \vect{v}^{\psi}_{p}\ h^{-1}\ m_p ,
  \end{aligned} 
\end{equation}
\end{itemize}
where $h$ is the thickness of the continuum in a 2D case. Here is
where the second discretization procedure appears. A background mesh
composed by a finite set of grid points with coordinates $\vec{x}_I\
, I = 1,2\ldots ,N_n$, is generated. Where $N_n$ is the number of grid
nodes. This mesh is employed as a support to compute gradients and divergences.

Introducing \eqref{eq:particle_acceleration_forces},
\eqref{eq:particle_internal_forces}, \eqref{eq:particle_body_forces}, 
and \eqref{eq:particle_load_forces} in \eqref{eq:BalanceMomentum_wf},
and approximating the real and virtual velocity  field of the particle
$p$ as
\begin{equation}
  \label{eq:Approx-velocity-mesh}
  \vec{v}_{p} = N_{Ip} \vect{v}_I
  \quad \text{and} \quad \vec{v}^{\psi}_{p} = N_{Ip} \vect{v}_I^{\psi},
\end{equation}
\begin{equation}
  \label{eq:Approx-gradvelocity-mesh}
  \tens{\dot{\varepsilon}}_p = (\vect{v}_I \otimes \Grad{N_{Ip}})^s
  \quad \text{and} \quad  \tens{\dot{\varepsilon}}_p^{\psi}=
  (\vect{v}_I^{\psi} \otimes \Grad{N_{Ip}})^s ,
\end{equation}
we reach to the particle balance of forces of the continuum,
\begin{equation}
  \label{eq:particle_balance_forces3}
  \dot{\vec{p}}_{I}= \tens{m}_{IJ}\dot{\vec{v}}_{J} = \vec{f}_{I}^{int} + \vec{f}_{I}^{ext},
\end{equation}
where $\dot{\vec{p}}_{I}$ is the rate of momentum at grid node $I$, the nodal mass matrix $\tens{m}_{IJ}$ is,
\begin{equation}
  \label{eq:particle_nod_mass}
  \tens{m}_{IJ} = N_{Ip} m_p N_{Jp}.
\end{equation}
To improve the computational efficiency and stability, the nodal mass matrix
\eqref{eq:particle_nod_mass} can be substituted by the lumped mass
matrix $\tens{m}_{IJ}^{lumped}$.
Later, internal and external forces are computed as follows,
\begin{equation}
  \label{eq:nodal_internal_forces}
  \vec{f}_{I}^{int} = - \tens{\sigma}_{p} \cdot \Grad{N_{Ip}} \frac{m_p}{\rho_p}
\end{equation}
\begin{equation}
  \label{eq:nodal_external_forces}
  \vec{f}_{I}^{ext} = N_{Ip}\ \vec{b}_{p}\ m_p  + N_{Ip}\ \vec{t}^s_{p}\ m_p h^{-1} 
\end{equation}
where $\tens{\sigma}_{p} = \tens{\sigma}_{p}(\tens{\varepsilon}_{p})$
is the particle $p$ stress field, which can be integrated employing
the suitable constitutive model. The strain tensor is updated employing the rate of stress tensor $\dot{ \tens{\varepsilon}}_{p}$ used to update the
strain tensor is as follows \eqref{eq:IncrStrainPoint}.
\begin{equation}
  \label{eq:IncrStrainPoint}
  \dot{\tens{\varepsilon}_{p}} = \frac{\Delta
    \tens{\varepsilon}_{p}}{\Delta t} =
  \frac{1}{2} \left[\Grad{N_{Ip}}\ \otimes \vec{v}_{I} + \vec{v}_{I} \otimes
    \Grad{N_{Ip}}\ \right].
\end{equation}
Next, imposing $\frac{D \rho}{D t} = 0$, we ensures the mass
conservation and update the density field.
\begin{equation}
  \label{eq:MassConservation}
\dot{\rho} = - \rho\ \mathit{tra} \left( \dot{\tens{\varepsilon}} \right)
\end{equation}

Finally, to solve the equation \eqref{eq:particle_balance_forces3}, a second order
temporal integration scheme is required. Therefore, time is
discretized in to a finite set of time steps $k = 1\ldots ,Nt$, where $k$ is the current time step and $N_t$
is the total number of time steps. Once the nodal equilibrium equation it is solved, the values in the nodes are
interpolated back in to the particles and each particle it is advected
to the new position,
\begin{equation}
  \label{eq:Updated_Lagrangian}
  \dot{\vec{v}}_p = N_{Ip}\ \vec{a}_{I},\quad and\quad
  \dot{\vec{x}}_{p} = N_{Ip}\ \vec{v}_{I}  
\end{equation}
In the MPM literature, this equations
(\ref{eq:particle_balance_forces3}) and (\ref{eq:Updated_Lagrangian}),
are solved with an explicit forward Euler algorithm. The time integration scheme of the MPM has been described in detail by many researchers
\cite{Sulsky1994}, \cite{Bardenhagen2002}, \cite{thesis_Andersen_2009} and summarized
in Figure \ref{fig:MPM_algorithm}.

\begin{figure}\sidecaption
  \centering
  \resizebox{0.8\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/MPM_scheme}
  }
  \caption{MPM standard algorithm.}
  \label{fig:MPM_algorithm}
\end{figure}

Other authors have proposed many others time integration alternatives
like \cite{Guilkey2003}, \cite{Tran2019e}, \cite{Charlton2017}. In the
first publication on the MPM \cite{Sulsky1994}, the nodal acceleration
was employed to update the particles as
\begin{equation}
  \label{eq:Sulsky-1994-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{a}_{I}^{k}
\end{equation}
\begin{equation}
  \label{eq:Sulsky-1994-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ N_{Ip}^{k}\ \vec{v}_{I}^{k}.
\end{equation}
However, as Andersen (2009)\cite{thesis_Andersen_2009} point out, this algorithm has been shown to be numerically unstable due to that
$\vect{f}_I^{int,k}$ can be finite for an infinitesimal nodal mass
$\tens{m}$. This can lead to numerical issues when nodal acceleration
is obtained for evaluating \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v}. Hence, a
corrected version of this algorithm in shown in Zhang {\it et al.}
(2016)\cite{Zhang_book_2016}
\begin{equation}
  \label{eq:Zhang-2016-UL-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{p}_{I}^{k}}{\tens{m}_I}.  
\end{equation}
\begin{equation}
  \label{eq:Zhang-2016-UL-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\ \frac{N_{Ip}^{k}\ \vec{f}_{I}^{k}}{\tens{m}_I},
\end{equation}
Later Tran \& Solowski (2019)\cite{Tran2019e} presented a
generalized-$\alpha$ scheme for the MPM inspired in the explicit time
integration algorithm proposed by Chung \& Hulbert
(1993)\cite{Geranlized_alpha_1993}, but with the particularity that
the acceleration is evaluated both in the beginning and the end of the
time step.
\begin{equation}
  \label{eq:Tran-2019-GA-v}
  \vect{v}_p^{k+1} = \vect{v}_p^{k} + \Delta t\  N_{Ip}^{k}\ \left[(1 - \gamma)\ \vect{a}_I^{k} +
    \gamma\ \vect{a}_I^{k+1} \right],\\
\end{equation}
\begin{equation}
\label{eq:Tran-2019-GA-x}
  \vect{x}_p^{k+1} = \vect{x}_p^{k} + N_{Ip}^{k} \left[ \Delta t\ \vec{v}_{I}^{k}+ \Delta t^2\left( (\frac{1}{2} - \beta)\
    \vec{a}_{I}^{k} + \beta\ \vec{a}_{I}^{k+1} \right) \right]
\end{equation}
\begin{equation}
  \label{eq:Tran-2019-GA-a}
  \vect{a}_p^{k+1} = N_{Ip}^{k}\ \vec{a}_{I}^{k+1}.
\end{equation}

This scheme has prof to damps out the higher frequency noises
\cite{Tran2019e}. But it can present the same numerical instabilities
as in \eqref{eq:Sulsky-1994-UL-x},\eqref{eq:Sulsky-1994-UL-v} when
nodal masses become infinitesimal, and requires extra storage for
nodal values of acceleration and previous steps.

@@

In this section, a explicit predictor-corrector time integration
scheme is presented. It is based in the Newmark a-form 
$\gamma = 0.5$ and $\beta = 0$ which is the central difference
explicit. This method is devoted to solve a system of equations of the type
\begin{equation*}
  \Matrix{M}_{IJ}\ddot{\Vector{d}}_{J} + \Matrix{C}_{IJ}\dot{\Vector{d}}_{J} +
  \Matrix{K}_{IJ}\Vector{d}_{J} = \Vector{F}_{I}.
\end{equation*}

As the MPM has a nodal stage, it is possible to apply this methods
successfully in the MPM framework as was proved by
\cite{Tran2019e}.

Taking the predictor definition from and calculating nodal velocity, and updating
particles position employing nodal values of velocity and acceleration.

The predictor-corrector algorithm has been described in the classic literature
\cite{Hughes2000}, and its computational advantages and stability were widely
proof by Liu \cite{Xiaojian94}. The ``classic'' PCE algorithm starts
with a predicted value of the nodal velocity at the $(k+1)$th time step denoted by $\vec{\tilde{v}}_I^{k+1}$ given as,
\begin{equation}
  \label{eq:Predictor-velocity-I}
  \vec{\tilde{v}}_I^{k+1} = \vec{v}_I^k + (1 - \gamma)\ \Delta t\ \vec{a}_I^k
\end{equation}

In \eqref{eq:Predictor-velocity-I} arise a \textit{user-defined}
parameter $\gamma \geq 0$. This parameter influences both the predictor accuracy
and the stability of the algorithm. As point out Liu
\cite{Xiaojian94}, the truncation error of the predictor formula is
$O(\Delta t^3)$ when $\gamma = 0.5$, and is unconditionally stable if
$ 0 < \gamma \leq 0.25$.
